---
title: "Geocomputation with R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://geocompr.robinlovelace.net/
https://github.com/Robinlovelace/geocompr 

```{r}
library(sf)          # classes and functions for vector data
library(raster)      # classes and functions for raster data
library(spData)      # load geographic data
library(sp)          # for use with spatial pckages which do not support sf data model


```

https://github.com/Robinlovelace/geocompr/issues/126
```{r}
# install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/", type = "source")

library(spDataLarge)   # load larger geographic data
```

```{r}
names(world)
```


```{r}
plot(world)
```

```{r}
summary(world["lifeExp"])
```

`sf` objects are easy to subset.
The code below shows its first two rows and three columns.
The output shows two major differences compared with a regular `data.frame`: the inclusion of additional geographic data (`geometry type`, `dimension`, `bbox` and CRS information - `epsg (SRID)`, `proj4string`), and the presence of final `geometry` column:

```{r}
# shows its first two rows and three columns

world[1:2, 1:3]
```

ome spatial packages (including **tmap**, **mapview** and **tidycensus**) have added support for **sf**.
However, it will take many years for most packages to transition and some will never switch.
Fortunately these can still be used in a workflow based on `sf` objects, by converting them to the `Spatial` class used in **sp**: 

```{r, eval=FALSE}
library(sp)
world_sp = as(world, Class = "Spatial")
# sp functions ...
```

`Spatial` objects can be converted back to `sf` in the same way or with `st_as_sf()`:

```{r, eval=FALSE}
world_sf = st_as_sf(world_sp, "sf")
```

### Basic map making {#basic-map}

Basic maps are created in **sf** with `plot()`.
By default this creates a multi-panel plot (like **sp**'s `spplot()`), one sub-plot for each variable of the object, as illustrated in the left-hand panel in Figure \@ref(fig:sfplot).
A legend or 'key' with a continuous color is produced if the object to be plotted has a single variable (see the right-hand panel).
Colors can also be set with `col = `, although this will not create a continuous palette or a legend. 

```{r sfplot, fig.cap="Plotting with sf, with multiple variables (left) and a single variable (right).", out.width="49%", fig.show='hold', warning=FALSE}
plot(world[3:4])
plot(world["pop"])
```

Plots are added as layers to existing images by setting `add = TRUE`.^[
`plot()`ing of **sf** objects uses `sf:::plot.sf()` behind the scenes.
`plot()` is a generic method that behaves differently depending on the class of object being plotted.
]
To demonstrate this, and to provide a taster of content covered in chapters \@ref(attr) and \@ref(spatial-operations) on attribute and spatial data operations, the subsequent code chunk combines countries in Asia:

```{r, warning=FALSE}
asia = world[world$continent == "Asia", ]
asia = st_union(asia)
```

We can now plot the Asian continent over a map of the world.
Note that the first plot must only have one facet for `add = TRUE` to work.
If the first plot has a key, `reset = FALSE` must be used:

```{r asia, out.width='50%', fig.cap="A plot of Asia added as a layer on top of countries worldwide."}
plot(world["pop"], reset = FALSE)
plot(asia, add = TRUE, col = "red")
```

Adding layers in this way can be used to verify the geographic correspondence between layers: 
the `plot()` function is fast to execute and requires few lines of code, but does not create interactive maps with a wide range of options.
For more advanced map making we recommend using dedicated visualization packages such as **tmap** (see Chapter \@ref(adv-map)).


```{r}
world_proj = st_transform(world, "+proj=eck4")
old_mar = par()$mar
par(mar = c(0, 0, 0, 0))
world_centroids_largest = st_centroid(world_proj, of_largest_polygon = TRUE)
main = "\nCountry continents and populations worldwide"
plot(st_geometry(world_proj), graticule = TRUE, reset = FALSE)
plot(world_proj["continent"], add = TRUE)
cex = sqrt(world$pop) / 10000
plot(st_geometry(world_centroids_largest), add = TRUE, cex = cex, lwd = 2)
par(mar = old_mar)
```


```{r}
plot(world["continent"], reset = FALSE)
cex = sqrt(world$pop) / 10000
plot(st_geometry(world_centroids), add = TRUE, cex = cex)
```
























