---
title: "Total 2016 Manufacturing Output, United States"
author: Scott Reeves
# date: 2018-05-20
output: pdf_document
    #toc: true
    #toc_float: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(comment = "#>",collapse = TRUE,cache = TRUE, fig.path = "images/",out.width = "100%",fig.align = 'center',fig.width = 10, fig.asp = 1)

```

## Sources and References

[Package ‘tmap’ .pdf](https://cran.r-project.org/web/packages/tmap/tmap.pdf)

[tmap in a nutshell](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html)

[Shapefiles - States](https://www.census.gov/geo/maps-data/data/cbf/cbf_state.html)

[About creating tmap insets for AK and HI](https://github.com/mtennekes/tmap/tree/master/demo/USChoropleth)


## Libraries

```{r message=FALSE, warning=FALSE}

library(tmap)
library(tmaptools)
library(grid)
library(sf)
library(tidyverse)
# library(leaflet)

```

```{r}
tidyverse_packages()
```

## Package 'tmap'

Author: Martijn Tennekes, etal

## tmap-package

Thematic maps are geographical maps in which spatial data distributions are visualized. This package offers a flexible, layer-based, and easy to use approach to create thematic maps, such as choropleths and bubble maps. It is based on the grammar of graphics, and resembles the syntax of `ggplot2`.

## Data

```{r}
mfg <- read.csv("F:/datasets/natlassocofmanufacturers/2017_state_mfg.csv")

states <- st_read("F:/datasets/uscensus/cb_2017_us_state_5m/cb_2017_us_state_5m.shp")
```

### Subset states census data

```{r}
outside51 <- c("Puerto Rico", "American Samoa", "United States Virgin Islands", "Guam", "Commonwealth of the Northern Mariana Islands")

states51 <- filter(states, !(NAME %in% outside51))
```

```{r eval=FALSE}
glimpse(states51)
```

```{r eval=FALSE}
# check to see if lengths are equal
length(mfg$state) == length(states51$NAME)
```

```{r eval=FALSE}
glimpse(mfg)
```

### Join spatial and mfg data into sf data frame

```{r}
mfg_map <- append_data(states51, mfg, key.shp = "NAME", key.data = "state")
```

### Further subset data for map insets

```{r}
US_cont <- mfg_map %>% 
  subset(!NAME %in% c("Alaska", "Hawaii")) %>% 
  simplify_shape(0.2) 

US_AK <- mfg_map %>% 
  subset(NAME == "Alaska") %>% 
  simplify_shape(0.2)

US_HI <- mfg_map %>% 
  subset(NAME == "Hawaii") %>% 
  simplify_shape(0.2) 

```

## Map 

```{r}
m_cont <- tm_shape(US_cont, projection = 2163) +
  tm_polygons("total_mfg_bil", 
              palette="YlOrRd",
              title="",
              border.col = "grey50",
              border.alpha = .5) +
  tm_credits("Data @ National Association of Manafacturers\nShape @ US Census Bureau", 
             position = c("right", "bottom")) +
  tm_layout(title="Total Manufacturing Output (in $billions, 2016)",
            title.position = c("center", "top"),
            legend.position = c("right", "bottom"),
            frame = FALSE,
            inner.margins = c(0.25, 0.15, 0.15, 0.15))



m_AK <- tm_shape(US_AK, projection = 3338) +
	tm_polygons("total_mfg_bil", 
	            palette="YlOrRd",
	            border.col = "gray50", 
	            border.alpha = .5, 
	            breaks = seq(10, 50, by = 5)) +
	tm_layout("Alaska", 
	          legend.show = FALSE, 
	          bg.color = NA, 
	          title.size = 0.8, 
	          frame = FALSE)



m_HI <- tm_shape(US_HI, projection = 3759) +
	tm_polygons("total_mfg_bil", 
	            palette="YlOrRd",
	            border.col = "gray50", 
	            border.alpha = .5, 
	            breaks=seq(10, 50, by = 5)) +
	tm_layout("Hawaii", 
	          legend.show = FALSE, 
	          bg.color=NA, 
	          title.position = c("LEFT", "BOTTOM"), 
	          title.size = 0.8, frame=FALSE)

```

### Set viewports

```{r}
vp_AK <- viewport(x = 0.15, y = 0.15, width = 0.3, height = 0.3)
vp_HI <- viewport(x = 0.4, y = 0.1, width = 0.2, height = 0.1)

```

### Print plot

```{r}
tmap_mode("plot")
m_cont
print(m_AK, vp = vp_AK)
print(m_HI, vp = vp_HI)
```

## Save image

```{r}
save_tmap(m_cont, "us_mfg_output_2016.png",
          scale = 0.7, width = 5.125,
          insets_tm = list(m_AK, m_HI),
          insets_vp = list(vp_AK, vp_HI))
```



















