---
title: "Mapping - Sandy Spring, GA"
author: "Scott Reeves"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

##  Notes and shortcut reminders

Add a new chunk:
by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

Run chunk:
by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Preview/Knit:
When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file). The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

Assignment:
  Alt/Opt+-

Commment/Uncomment:
  Ctrl/Cmd+Shift+C

## Data Sources

Open Data, City of Sandy Springs
https://data-coss.opendata.arcgis.com/

## Load Library Packages

```{r message=FALSE, warning=FALSE}
library(tmap)
library(tmaptools)
library(rio)
library(tidyverse)
```

## Load Data

```{r}
# Assign fiile paths to vars

districtsshapefile <- "F:/datasets/sandyspringsga/City_Council__Sandy_Springs_Georgia/City_Council__Sandy_Springs_Georgia.shp"

crimedatafile <- "F:/datasets/sandyspringsga/Sandy_Springs_Police_Department.xlsx"
```

```{r}
# Assign shapefile to a variable using tmaptools's read_shape()
# as.sf = TRUE means as a simple features object (sf package) akin to a regualr dataframe.

ssgeo <- read_shape(districtsshapefile, as.sf = TRUE)

```

```{r}
# Try a couple of quick basic plots

plot(ssgeo)
```

```{r}
qtm(ssgeo)
```

```{r}
qtm(ssgeo, fill = "DISTRICT")
```


```{r}
# Assign data file to var
sscrime <- rio::import(crimedatafile)
```

```{r}
#  View the shapefile structure
str(ssgeo)
```

```{r}
# View datafile structure
str(sscrime)
```
```{r}
# Subset the crime data
sscrime2018 <- dplyr::filter(sscrime, incident_datetime > "2017-12-31")
```

```{r}
tm_shape(ssgeo) +
tm_dots(sscrime2018)

```

```{r}
map1 + tm_pol
```

