---
title: "tmap"
author: Scott Reeves
# date: 2018-05-20
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
# set.seed(1014)
# options(digits = 3)

knitr::opts_chunk$set(comment = "#>",collapse = TRUE,cache = TRUE, fig.path = "images/",out.width = "100%",fig.align = 'center',fig.width = 10, fig.asp = 1)

# options(dplyr.print_min = 5, dplyr.print_max = 5)
```

# Sources and References

[Package ‘tmap’ .pdf](https://cran.r-project.org/web/packages/tmap/tmap.pdf)

[tmap in a nutshell](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html)

[Shapefiles - States](https://www.census.gov/geo/maps-data/data/cbf/cbf_state.html)

[About creating tmap insets for AK and HI](https://github.com/mtennekes/tmap/tree/master/demo/USChoropleth)

[Leaflet Providers](http://leaflet-extras.github.io/leaflet-providers/preview/)



# Libraries

```{r message=FALSE, warning=FALSE}

library(tmap)
library(tmaptools)
library(grid)
library(sf)
library(tidyverse)
library(leaflet)

```

```{r}
tidyverse_packages()
```

# Package 'tmap'

Author: Martijn Tennekes, etal

## tmap-package

Thematic maps are geographical maps in which spatial data distributions are visualized. This package offers a flexible, layer-based, and easy to use approach to create thematic maps, such as choropleths and bubble maps. It is based on the grammar of graphics, and resembles the syntax of `ggplot2`.

## package functions

### Quick plotting method - qtm

Plot a thematic map

### Main plotting method - tm_shape

Specify a shape object

Aesthetics base layers: | 
---|---
`tm_polygons` | Create a polygon layer (with borders)
`tm_symbols` | Create a layer of symbols
`tm_lines` | Create a layer of lines
`tm_raster` | Create a raster layer
`tm_text` | Create a layer of text labels

Aesthetics derived layers: |  
---|---
`tm_fill` | Create a polygon layer (without borders)
`tm_borders` | Create polygon borders
`tm_bubbles` | Create a layer of bubbles
`tm_squares` | Create a layer of squares
`tm_dots` | Create a layer of dots
`tm_markers` | Create a layer of markers
`tm_iso` | Create a iso/contour lines

More - see [tmap.pdf]

# Data

Spatial datasets (included in tmap package) |
---|---
World World | country data (spatial polygons)
Europe European | country data (spatial polygons)
NLD_prov | Netherlands province data (spatial polygons)
NLD_muni | Netherlands municipal data (spatial polygons)
metro | Metropolitan areas (spatial points)
rivers | Rivers (spatial lines)
land | Global land cover (spatial grid)

```{r}
# from tmap package
data(World, metro, rivers)
```


```{r}
# convert to sf objects
metro_sf <- st_as_sf(metro)
World_sf <- st_as_sf(World)

```

## World, Europe, and the Netherlands

Maps of the world, Europe, and the Netherlands (province and municipality level), class `SpatialPolygonsDataFrame`

The default projections for these maps are Eckhart IV (World), Lambert azimuthal (Europe), and Rijksdriehoekstelsel (Netherlands). See below. To change the projection, use set_projection. Alternatively, the projection can be changed temporarily for plotting purposes by using the projection
argument of `tm_shape` (or `qtm`).

World World map. The default projection for this world map is Eckhart IV since area sizes are preserved, which is a very important property for choropleths.

Europe Europe map. The ETRS-LAEA projection is used by default for this map. This projection, as well as the bounding box, are also used in the maps published by Eurostat. Several countries are transcontinental and are partly located in Asia. From these countries, only data from Russia and Turkey have been included in this shape file since they are widely considered as European countries. The data is taken from the World data, where variables "part" and "EU_Schengen" have been added.

NLD_prov and NLD_muni, maps of the Netherlands at province and municipality level of 2013. The used projection is the Rijksdriehoekstelsel projection. Important: publication of these maps is only allowed when cited to Statistics Netherlands (CBS) and Kadaster Nederland as source.


## metro

Spatial data of metropolitan areas, of class SpatialPointsDataFrame. The data includes a population times series from 1950 to (forecasted) 2030. All metro areas with over 1 million inhabitants in 2010 are included.

```{r}
glimpse(metro_sf)
```

## rivers

Spatial data of rivers, of class `SpatialLinesDataFrame`

Source
http://www.naturalearthdata.com


## land

Spatial data of global land cover, of class `SpatialGridDataFrame`. The data includes a population times series from 1950 to (forecasted) 2030. All metro areas with over 1 million inhabitants in 2010 are included.


# animation_tmap

_Interesting concept/element. Revisit later. _

Create a gif or mpeg animation from a tmap plot. The free tool ImageMagick is required.

# qtm

__Description__

Draw a thematic map quickly.

__Usage__

qtm(shp, fill = NA, symbols.size = NULL, symbols.col = NULL,
symbols.shape = NULL, dots.col = NULL, text = NULL, text.size = 1,
text.col = NA, lines.lwd = NULL, lines.col = NULL, raster = NA,
borders = NA, by = NULL, scale = NA, title = NA, projection = NULL,
format = NULL, style = NULL, basemaps = NA, bubble.size = NULL,
bubble.col = NULL, ...)

__Arguments__

See documentation.

## just the map
```{r}
qtm(World) #SpatialPolygons shape object. 

```

## choropleth
```{r}
qtm(World, 
    fill = "economy", # fill the polygons based on data (other option would be to specify a color)
    format="World", # tm_layout wrapper used for format.1 of 6 currently available in tmap. 
    style="col_blind") # tm_layout wrapper used for style.
```

```{r}
qtm(World, fill="HPI", fill.n=9, fill.palette="div", fill.auto.palette.mapping=FALSE,
fill.title="Happy Planet Index", fill.id="name", format="World", style="gray")

```

## bubble map
```{r}
current.mode <- tmap_mode("view")

qtm(World, borders = NULL) +
qtm(metro, symbols.size = "pop2010",
symbols.title.size= "Metropolitan Areas",
symbols.id= "name",
format = "World")
```

# tmap-element

__Description__

Building block for drawing thematic maps. All element functions have the prefix tm_.

__Details__

The fundamental, and hence required element is 
* `tm_shape` that specifies the shape object, and also specifies the projection and bounding box
The elements that serve as aesthetics layers are
* `tm_fil`l to fill the polygons
* `tm_borders` to draw polygon borders
* tm_polygons to draw polygons (it is a combination of tm_fill and tm_borders)
* `tm_bubbles` to draw bubbles
* `tm_lines` to draw lines
* `tm_text` to add text annotations
* `tm_raster` to draw a raster

The layers can be stacked by simply adding them with the + symbol. The combination of the elements described above form one group. Multiple groups can be stacked. Each group should start with tm_shape.

The attribute elements are
* `tm_grid` to specify coordinate grid lines
* `tm_credits` to add a credits/acknowledgements text label
* `tm_scale_bar` to add a measurement scale bar
* `tm_compass` to add a map compass

The element `tm_facets` specifies facets (small multiples). The element `tm_layout` is used to change the layout of the map.

## tmap_arrange

FILL LATER

## tmap_icons

FILL LATER

## tmap_leaflet 

__Description__

Create a leaflet widget from a tmap object. An interactive map (see tmap_mode) is an automatically generated leaflet widget. With this function, this leaflet widget is obtained, which can then be changed or extended by using leaflet’s own methods.

FILL LATER

## tmap_mode

FILL LATER

## tmap_options

FILL LATER

## tm_add_legend

FILL LATER

## tm_compass 

FILL LATER

## tm_credits ETC - FILL LATER ...

## Draw Polygons - tm_fill, tm_borders, tm_polygons

__Description__

Creates a `tmap-element` that draws the polygons. `tm_fill` fills the polygons. Either a fixed color is used, or a color palette is mapped to a data variable. `tm_borders` draws the borders of the polygons. `tm_polygons` fills the polygons and draws the polygon borders.


### tm_fill

### tm_borders

### tm_polygons





## tm_shape

Specify the shape object

__Description___

Creates a `tmap-element` that specifies the shape object. Also the projection and covered area (bounding box) can be set. It is possible to use multiple shape objects within one plot (see tmap-element).

__Usage___

tm_shape(shp, name = NULL, is.master = NA, projection = NULL,
bbox = NULL, unit = getOption("tmap.unit"), simplify = 1,
line.center.type = c("segment", "midpoint"), ...)

__Arguments___

shp shape object, which is one of:

1. SpatialPolygons(DataFrame)
2. SpatialPoints(DataFrame)
3. SpatialLines(DataFrame)
4. SpatialGrid(DataFrame)
5. SpatialPixels(DataFrame)
6. RasterLayer, RasterStack, or RasterBrick

Simple features (sf objects) are also supported. For drawing layers `tm_fill` and `tm_borders`, 1 is required. For drawing layer `tm_lines`, 3 is required. Layers `tm_symbols` and `tm_text` accept 1 to 3. For layer `tm_raster`, 4, 5, or 6 is required.

* More arguments, see documentation.

https://cran.r-project.org/web/packages/tmap/vignettes/tmap-nutshell.html

The first element to start with is `tm_shape`, which specifies the shape object. Next, one, or a combination of the following drawing layers should be specified:

Drawing layer | Description | Aesthetics
---|---|---
Base layer | |
`tm_polygons` | Draw polygons	| col
`tm_symbols` | Draws symbols | size, col, shape
`tm_lines` | Draws polylines | col, lwd
`tm_raster` | Draws a raster | col
`tm_text` | Add text labels | text, size, col
Derived layers |  | 
`tm_fill` | Fills the polygons | see tm_polygons
`tm_borders` | Draws polygon borders | none
`tm_bubbles` | Draws bubbles | see tm_symbols
`tm_squares` | Draws squares | see tm_symbols
`tm_dots` | Draws dots | see tm_symbols
`tm_markers` | Draws markers | see tm_symbols and tm_text
`tm_iso` | Draws iso/contour lines | see tm_lines and tm_text




```{r message=FALSE}

current.mode <- tmap_mode("plot")


tm_shape(World, projection="longlat") +
tm_polygons() +
tm_layout("Long lat coordinates (WGS84)", inner.margins=c(0,0,.1,0), title.size=.8)

```


```{r}
#from save_tmap

current.mode <- tmap_mode("plot")

data(NLD_muni, NLD_prov)

m <- tm_shape(NLD_muni) +
  tm_fill(col="population", convert2density=TRUE, style="kmeans", title=expression("Population (per " * km^2 * ")"), legend.hist=FALSE) +
  tm_borders("black", alpha=.5) + 
  tm_shape(NLD_prov) + 
  tm_borders("grey25", lwd=2) + 
  tm_format_NLD(inner.margins = c(.02, .15, .06, .15)) + 
  tm_scale_bar(position = c("left", "bottom")) +
  tm_compass(position=c("right", "bottom")) +
  tm_style_classic()
m

```


# Worflow with sf and tmap

## Data

Read in data

```{r}
mfg <- read.csv("F:/datasets/natlassocofmanufacturers/2017_state_mfg.csv")

states <- st_read("F:/datasets/uscensus/cb_2017_us_state_5m/cb_2017_us_state_5m.shp")
```



### Cleanse data


```{r}
outside51 <- c("Puerto Rico", "American Samoa", "United States Virgin Islands", "Guam", "Commonwealth of the Northern Mariana Islands")

states51 <- filter(states, !(NAME %in% outside51))
```

```{r eval=FALSE}
glimpse(states51)
```

```{r}
# check to see if lengths are equal
length(mfg$state) == length(states51$NAME)
```

```{r eval=FALSE}
glimpse(mfg)
```


### Join data

```{r}
mfg_map <- append_data(states51, mfg, key.shp = "NAME", key.data = "state")
```

```{r}
US_cont <- mfg_map %>% 
  subset(!NAME %in% c("Alaska", "Hawaii")) %>% 
  simplify_shape(0.2) 

US_AK <- mfg_map %>% 
  subset(NAME == "Alaska") %>% 
  simplify_shape(0.2)

US_HI <- mfg_map %>% 
  subset(NAME == "Hawaii") %>% 
  simplify_shape(0.2) 

```

## Map data


```{r}
m_cont <- tm_shape(US_cont, projection = 2163) +
  tm_polygons("total_mfg_bil", 
              palette="Reds",
              title="",
              border.col = "grey50",
              border.alpha = .5) +
  tm_credits("Data @ National Association of Manafacturers\nShape @ US Census Bureau", 
             position = c("right", "bottom")) +
  tm_layout(title="Total Manufacturing Output (in $billions, 2016)",
            title.position = c("center", "top"),
            legend.position = c("right", "bottom"),
            frame = FALSE,
            inner.margins = c(0.25, 0.15, 0.15, 0.15))



m_AK <- tm_shape(US_AK, projection = 3338) +
	tm_polygons("total_mfg_bil", 
	            palette="Reds",
	            border.col = "gray50", 
	            border.alpha = .5, 
	            breaks = seq(10, 50, by = 5)) +
	tm_layout("Alaska", 
	          legend.show = FALSE, 
	          bg.color = NA, 
	          title.size = 0.8, 
	          frame = FALSE)



m_HI <- tm_shape(US_HI, projection = 3759) +
	tm_polygons("total_mfg_bil", 
	            palette="Reds",
	            border.col = "gray50", 
	            border.alpha = .5, 
	            breaks=seq(10, 50, by = 5)) +
	tm_layout("Hawaii", 
	          legend.show = FALSE, 
	          bg.color=NA, 
	          title.position = c("LEFT", "BOTTOM"), 
	          title.size = 0.8, frame=FALSE)

```


```{r}
vp_AK <- viewport(x = 0.15, y = 0.15, width = 0.3, height = 0.3)
vp_HI <- viewport(x = 0.4, y = 0.1, width = 0.2, height = 0.1)

```


```{r}
tmap_mode("plot")
m_cont
print(m_AK, vp = vp_AK)
print(m_HI, vp = vp_HI)
```


```{r}
save_tmap(m_cont, "us_mfg_output_2016.png",
          insets_tm = list(m_AK, m_HI),
          insets_vp = list(vp_AK, vp_HI))
```



















